Cassini State Capture

### 2019-10-10 writeups/Su_disk_disp.tex
stash draft for 10/10 meeting

@@

Denote $\vec{s}$ spin of planet, $\vec{l}$ angular momentum of planet, and
$\vec{l}_d$ angular momentum of the surrounding disc. We approximate $S \ll L
\ll L_d$, so $\vec{l}_d$ is approximately constant. Much of this treatment
borrows from \citep{anderson2018teeter}. We consider precession equations
\begin{align}
    \rd{\hat{s}}{t} &= \omega_{sl} \p{\hat{s} \cdot \hat{l}}
        \p{\hat{s} \times \hat{l}}\label{eq:dsdt}\\
    \rd{\hat{l}}{t} &= \omega_{ld}\p{\hat{l} \cdot \hat{l}_d}
        \p{\hat{l} \times \hat{l}_d},\label{eq:dldt},\\
\end{align}
where
\begin{align}
    \omega_{sl} &\equiv \frac{3k_{qp}}{2k_p} \frac{M_\star}{m_p}
        \p{\frac{R_p}{a_p}}^3 \Omega_p,\label{eq:wsl}\\
    \omega_{ld} &\equiv \frac{3M_d}{4M_\star}\p{\frac{a_p}{r_d}}^3 n
        .\label{eq:wld}
\end{align}
In \autoref{eq:wsl}, $k_{qp}$ and $k_p$ are related to the planet's quadrupole
moment and moment of inertia respectively \citep[see][]{lai2018}, $M_\star$ is
the mass of the host star, and $M_p, R_p, a_p, \Omega_p$ are the mass, radius,
semi-major axis and angular spin frequency of the planet respectively.
\autoref{eq:wld} is written for a disk of mass $M_d$ located at characteristic
distance $r_d$ from the host star \citep[see][for a power-law disk
profile]{millholland_disk}, $n \equiv \sqrt{GM_\star/a_p^3}$ is the planet's
orbital mean motion. The three angles formed by the three vectors are notated
\begin{align}
    \theta &\equiv \arccos \hat{s} \cdot \hat{l},\\
    \theta_{sd} &\equiv \arccos \hat{s} \cdot \hat{l}_d,\\
    I &\equiv \arccos \hat{l} \cdot \hat{l}_d.
\end{align}

@@

Spin states satisfying $\rd{\hat{s}}{t} = 0$ are referred to as \emph{Cassini
States} (CS). When $\eta < \eta_c$, there are four CSs, and when $\eta > \eta_c$
there are only two; $\eta_c$ is \citep{henrard1987,ward2004I}
\begin{equation}
    \eta_c \equiv \p{\sin^{2/3}I + \cos^{2/3}I}^{3/2}.
\end{equation}
CSs 1, 2, 3 are stable while CS4 is unstable. A contour plot of $H\p{\phi, \cos
\theta}$ is provided in \autoref{fig:eq_1contours}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/0_eta/1contours_flip.png}
    \caption{Contour plot of $H\p{\phi, \cos \theta}$, where warmer colors
    denote more positive values. The black solid line is the separatrix, which
    only exists for $\eta < \eta_c$. The three zones, divided by the separatrix,
    are labeled. The interior of the separatrix, shaded in grey, is formally
    only defined for $\eta < \eta_c$, but we may identify the phase space
    trajectories that flow into zone II when evolved forward in time; this is
    the shaded region in the top left plot, bounded by the dotted black
    line.}\label{fig:eq_1contours}
\end{figure}

@@

In the four-CS regime, one of the CSs is a saddle point, conventionally denoted
Cassini State 4 (CS4). All trajectories are periodic with finite period except
two critical trajectories asymptotic in the past and future to CS4. Together,
these two critical trajectories are referred to as the \emph{separatrix} and
divide phase space into three zones. We label these zones as illustrated in
\autoref{fig:eq_1contours}. As trajectories of constant $\eta$ evolve along
contours of constant $H$, it can be observed that trajectories in zone II
librate about CS2 while those in zones I/III circulate.

@@

\subsection{Fiducial Simulation: Zone Transition Histories}\label{ss:ad_fid}

@@

Initially, in the $\eta > \eta_c$ regime, only zones $II, III$ exist.
Conversely, at the end of the simulation when $\eta \to 0$, only zones $I, III$
exist. Naively then, one might expect four sequences of transitions
between zones (we call these dynamical ``tracks'') to appear during simulations:
(i) $II \to I$, (ii) $II \to III$, (iii) $III \to I$, and (iv) the trivial $III
\to III$ where the trajectory never encounters the separatrix. In fact, (v) a
fifth possible track involving two transitions is observed $III \to II \to I$.

Between zone transitions, the enclosed phase space area $A \equiv \oint \cos\theta
\;\mathrm{d}\phi$ is an adiabatic invariant. Here, we adopt convention $A$ is a
\emph{signed} area
\begin{equation}
    A = \oint \p{1 - \cos \theta}\;\mathrm{d}\phi.
\end{equation}
This has the advantage of (i) having a removable singularity when the trajectory
crosses $\cos \theta = 1$, and (ii) being easily expressible as combinations of
the $A_i$ of \autoref{se:area_ward}. The bounds of the integral are either a
libration or circulation cycle.

Given this definition, the change in $A$ during a separatrix encounter is easily
understood \citep{henrard1982}. We present in \autoref{fig:ad_21} a fiducial
simulation undergoing the $II \to I$ zone transition. An initial condition in
zone II with initial adiabatic invariant $A_i > 0$ librates about CS2 until
$A_{II} = A_i$. At this point, it encounters the separatrix and moves into zone
I and the final enclosed phase space area is $A_f = -(A_{II} + A_{III})$.
Further simulations illustrating the other tracks are depicted in
\autoref{ss:app_tracks}.
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/3testo21.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/3testo21_subplots.png}
    \end{subfigure}
    \caption{Fiducial simulation following the $A_{II} \to A_{I}$ transition. Top
    plot upper panel: plot of $\cos \theta(t)$ (gold) in an example simulation.
    Overlaid are the locations of Cassini State 2 (magenta), and upper and lower
    bounds on the separatrix (dashed black). Note that the trajectory
    successfully tracks CS2 to a misalignment angle $\theta \approx 90^\circ$.
    Vertical green lines denote times portrayed in bottom plot. Top plot lower
    panel: plot of the enclosed separatrix area obtained by integrating the
    simulated trajectory (blue dots) and adiabatic theory (red line). Lower
    plot: snapshots in the $\p{\cos \theta, \phi}$ space of the simulation
    trajectories for the times demarcated in green above, corresponding to the
    start of the simulation, the appearance of the separatrix, two panels
    depicting the separatrix crossing process, and a final snapshot at $\eta =
    10^{-3.5}$. Gold dots denote a full circulation or libration cycle at the
    selected time, while the separatrix at the start/end of the portrayed cycle
    are shown in solid/dashed black lines respectively. Also labeled is CS2
    (magenta).}\label{fig:ad_21}
\end{figure}

@@

In the adiabatic limit, while separatrix encounters induce discontinuous changes
in enclosed phase space, the outcomes of these encounters are well understood in
pioneering work by Henrard \citep{henrard1982,henrard1987}. For a wide range of
$\theta_{sd, i}$, we run simulations over a ring of initial conditions at each
$\theta_{sd, i}$ and measure the final $\theta_{ f}$ values. We can also
predict the values of $\theta_{ f}$ for each of the tracks as
well as their associated probabilities; refer to \autoref{ss:app_transition} for
details. The comparison of the final obliquities associated with these tracks
and those obtained via simulation can be seen in \autoref{fig:ad_ensemble}. The
same plot for $I = 10^\circ, I = 20^\circ$ can be found in
\autoref{ss:varying_I}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_05_35.png}
    \caption{Top: $\theta_{f}\p{\theta_{sd, i}}$, overlaid with semi-analytic
    predictions of the $\theta_{ f}$ for each of the four nontrivial
    dynamical tracks in colored lines. Bottom: Semi-analytic probabilities of
    each of the dynamical tracks for each $\theta_{sd,
    i}$.}\label{fig:ad_ensemble}
\end{figure}

@@

To illustrate the transition to non-adiabaticity, we present three simulations
with increasingly larger values of $\epsilon$ than \autoref{fig:ad_ensemble}
below in \autoref{fig:transition_to_nonad}.

@@

Later, these large scatters take on band-like structures that seem to persist
across values of $\theta_{sd, i}$; we attribute these to the separatrix crossing
process becoming sensitive to the \emph{phase} of the final
libration/circulation orbit at separatrix crossing. This is equivalent to
separatrix crossing no longer being significantly slower than orbit timescales
and violating the adiabatic assumption.
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_05_25.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_05_15.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_05_10.png}
    \end{subfigure}
    \caption{Increasingly larger $\epsilon$ used compared to those in
    \autoref{fig:ad_ensemble}. The general evolutionary tracks in the adiabatic
    limit deteriorate as $\epsilon$ is increased. Deviations from the adiabatic
    predictions seem to be banded structures, which can be attributed to
    non-adiabaticity ``freezing-in'' the phase of the obliquity variations over
    the final libration/circulation orbit prior to separatrix
    crossing.}\label{fig:transition_to_nonad}
\end{figure}

\section{Nonadiabatic Evolution}\label{s:nonad}

Here, we consider $\epsilon \sim 0.2$ which significantly violates
\autoref{eq:ad_constr}.

\subsection{Sample Trajectory}

@@

The agreement of \autoref{eq:nonad_q_f} at fixed $I$ for varying $\epsilon$ is
shown in \autoref{fig:nonad_3_scan}. Note that $\epsilon \to 0$ recovers the
adiabatic regime.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3scan.png}
    \caption{Plot of $\theta_{ f}\p{\theta_{sd, i} = 0}$ as a function of
    $\epsilon$, where $I = 5^\circ$. Overplotted in the red line is
    \autoref{eq:nonad_q_f}, which is in good agreement for $\epsilon \gtrsim
    0.1$ the non-adiabatic regime, while $\theta_{ f} \approx 90^\circ$ in
    the adiabatic regime.}\label{fig:nonad_3_scan}
\end{figure}

@@

\subsection{Adiabatic Dynamical Histories}\label{ss:app_tracks}

>>>>

% TODO
% figure out whether can draw vertical lines on 3testo plots for regimes
% run simulations up to III -> III (not today)

@@

Denote $\vec{s}$ spin of planet, $\vec{l}$ angular momentum of planet, and
$\vec{l}_d$ angular momentum of the surrounding disc. We approximate $S \ll L
\ll L_d$, so $\vec{l}_d$ is approximately constant. Much of this treatment
borrows from \citealp{anderson2018teeter} and \citealp{millholland_disk}. We
consider precession equations
\begin{align}
    \rd{\hat{s}}{t} &= \omega_{sl} \p{\hat{s} \cdot \hat{l}}
        \p{\hat{s} \times \hat{l}}\label{eq:dsdt}\\
    \rd{\hat{l}}{t} &= \omega_{ld}\p{\hat{l} \cdot \hat{l}_d}
        \p{\hat{l} \times \hat{l}_d},\label{eq:dldt}\\
\end{align}
where
\begin{align}
    \omega_{sl} &\equiv \frac{3k_{qp}}{2k_p} \frac{M_\star}{m_p}
        \p{\frac{R_p}{a_p}}^3 \Omega_p,\label{eq:wsl}\\
    \omega_{ld} &\equiv \frac{3M_d}{4M_\star}\p{\frac{a_p}{r_d}}^3 n
        .\label{eq:wld}
\end{align}
In \autoref{eq:wsl}, $k_{qp}$ and $k_p$ are related to the planet's quadrupole
moment and moment of inertia respectively \citep[see][]{lai2018}, $M_\star$ is
the mass of the host star, and $M_p, R_p, a_p, \Omega_p$ are the mass, radius,
semi-major axis and angular spin frequency of the planet respectively.
\autoref{eq:wld} is written for a disk of mass $M_d$ located at characteristic
distance $r_d$ from the host star \citep[see][for a power-law disk
profile]{millholland_disk}, $n \equiv \sqrt{GM_\star/a_p^3}$ is the planet's
orbital mean motion. The three angles formed by the three vectors are notated
\begin{align}
    \theta &\equiv \arccos \hat{s} \cdot \hat{l},\\
    \theta_{sd} &\equiv \arccos \hat{s} \cdot \hat{l}_d,\\
    I &\equiv \arccos \hat{l} \cdot \hat{l}_d.
\end{align}

@@

Spin states satisfying $\rd{\hat{s}}{t} = 0$ are referred to as \emph{Cassini
States} (CS). When $\eta < \eta_c$, there are four CSs, and when $\eta > \eta_c$
there are only two; $\eta_c$ is \citep{henrard1987,ward2004I}
\begin{equation}
    \eta_c \equiv \p{\sin^{2/3}I + \cos^{2/3}I}^{3/2}.
\end{equation}
CSs 1, 2, 3 are stable while CS4 is unstable.

@@

In the four-CS regime, one of the CSs is a saddle point, conventionally denoted
Cassini State 4 (CS4). All trajectories are periodic with finite period except
two critical trajectories asymptotic in the past and future to CS4. Together,
these two critical trajectories are referred to as the \emph{separatrix} and
divide phase space into three zones. The separatrix and three zones are most
clearly shown via a contour plot of $H\p{\phi, \cos \theta}$ in
\autoref{fig:eq_1contours}.
\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{../initial/0_eta/1contours_flip.png}
    \caption{Contour plot of $H\p{\phi, \cos \theta}$, where warmer colors
    denote more positive values. The black solid line is the separatrix, which
    only exists for $\eta < \eta_c$. The three zones, divided by the separatrix,
    are labeled. The interior of the separatrix, shaded in grey, is formally
    only defined for $\eta < \eta_c$, but we may identify the phase space
    trajectories that flow into zone II when evolved forward in time; this is
    the shaded region in the top left plot, bounded by the dotted black
    line.}\label{fig:eq_1contours}
\end{figure*}

As trajectories of constant $\eta$ evolve along
contours of constant $H$, it can be observed that trajectories in zone II
librate about CS2 while those in zones I/III circulate.

@@

\subsection{Separatrix Crossing and Evolutionary Tracks
}\label{ss:zone_transitions}

@@

In the absence of separatrix encounters, the enclosed phase space area $A \equiv
\oint \cos\theta \;\mathrm{d}\phi$ is an adiabatic invariant. Here, we adopt
convention where $A$ is a \emph{signed} area
\begin{equation}
    A = \oint \p{1 - \cos \theta}\;\mathrm{d}\phi.
\end{equation}
This has the advantage of (i) having a removable singularity when the trajectory
crosses $\cos \theta = 1$, and (ii) being easily expressible as combinations of
the $A_i$ of \autoref{se:area_ward}. The bounds of the integral are either a
libration ($\phi$ returns to its original value with the same $\dot{q\phi}$
sign) or circulation cycle ($\phi$ advances to $\phi \pm 2\pi$). Note lastly
that $A$ has simple physical interpretation of are enclosed on the unit sphere
by $\hat{s}$ over one cycle measured relative to $+\hat{z}$:
\begin{equation*}
    \int \int\limits_{\cos \theta(\phi)}^1
        \mathrm{d}\cos\theta\;\mathrm{d}\phi = \int \p{1 - \cos \theta(\phi)}
            \;\mathrm{d}\phi.
\end{equation*}

When a trajectory encounters the separatrix, $A$ will change discontinuously as
the trajectory moves between zones, but its change is easily understood
\citep{henrard1982}. We can classify trajectories by the separatrix encounters
they undergo. Initially, in the $\eta > \eta_c$ regime, only zones $II, III$
exist. Conversely, at the end of the simulation when $\eta \to 0$, only zones
$I, III$ exist. Most simply, one might expect four sequences of transitions
between zones (we call these dynamical ``tracks'') to manifest: from one of
$\z{I, III}$ to one of $\z{II, III}$. In addition to these four tracks, a
fifth track $III \to II \to I$ is observed.

Beyond simply identifying the possible tracks, we may clearly delineate the
initial conditions for which each track can be encountered. Separatrix crossing
is a probabilistic process whose outcomes are well understood by
\citealp{henrard1982,henrard1987}. Their resultss may be simply summarized as
follows: if a zone $i$ is shrinking and zone $j$ is expanding, the probability
of transition from zone $i$ to zone $j$ is given
\begin{equation}
    \Pr_{i \Rightarrow j} = -\frac{\pd{A_j}{t}}{ \pd{A_i}{t}}
        = -\frac{\pd{A_j}{\eta}}{ \pd{A_i}{\eta}}\label{eq:henrard_hop}
\end{equation}
This can be used directly in conjunction with \autoref{se:area_ward} to
understand for what initial conditions each track can be observed and with what
probabilities.

Below, we will describe the evolution of $A$ throughout each of the five tracks,
as well as determine the initial conditions and probabilities associated with
each track:
\begin{itemize}
    \item $II \to I$ --- This track is depicted in \autoref{fig:ad_21}.
        $\hat{s}$ starts librating about CS2 in zone II, enclosing some initial
        phase space area $A_i$. The trajectory is able to librate without
        separatrix encounter until $A_{II}(\eta_\star) = A_i$. At this point, as
        the trajectory transitions to a circulating trajectory in zone I
        immediately bordering the separatrix, it will encompass
        $-A_I(\eta_\star)$ phase space area. Conservation of phase space area
        then predicts that as $\eta \to 0$, the final enclosed phase space area
        will still be $A_f = -A_I$, and the final obliquity $\theta_f$ must be
        given by
        \begin{equation}
            2\pi\p{1 - \cos \theta_f} = A_f. \label{eq:qfaf}
        \end{equation}

        This transition can only occur when the initial condition begins in zone
        II, requiring $A_i < A_{II}(\eta_c)$, where $A_{II}(\eta_c)$ is known in
        closed form \citep{ward2004I}
        \begin{equation}
            A_{II}\p{\eta_c} = 4\pi \s{
                1 - \p{1 + \tan^{2/3}I}^{-3/2}}.
        \end{equation}
        Then, at some known $\eta_\star$ satisfying $A_{II}(\eta_\star) = A_i$,
        the trajectory is ejected from the separatrix and enters zone I with
        probability
        \begin{equation}
            \Pr_{II \to I} = -\at{\frac{\pd{A_I}{\eta}}{\pd{A_{II}}{\eta}}}
                _{\eta_\star}.
        \end{equation}
        Note that since $\pd{A_i}{\eta} < 0$ everywhere, while
        $\pd{A_{II}}{\eta} > 0$ at all possible $\eta_\star$ for an initial
        condition starting in zone II, this track always has nonzero
        probability.

    \item $II \to III$ --- This track is depicted in \autoref{fig:ad_23}. The
        only difference from the previous track is that, upon separatrix
        encounter, the trajectory follows the circulating trajectory in zone III
        bordering the separatrix, upon which it will encompass $A_f =
        A_I(\eta_\star) + A_{II}(\eta_\star)$. The final obliquity is still
        given by \autoref{eq:qfaf}.

        Again, this track can only occur when $A_i < A_{II}(\eta_c)$, and enters
        zone III with probability
        \begin{equation}
            \Pr_{II \to III} = -\at{\frac{\pd{A_{III}}{\eta}}
                {\pd{A_{II}}{\eta}}} _{\eta_\star}.
        \end{equation}
        Upon examination of \autoref{fig:eq_areas}, it is clear that
        $\pd{A_{III}}{\eta} > 0$ for many $\eta$. Call
        \begin{equation}
            \eta_{\min, III} \equiv \argmin A_{III}(\eta),
        \end{equation}
        then if $\eta_\star > \eta_{\min, III}$ then $\Pr_{II \to III} < 0$.
        This is interpreted as a forbidden transition, and so $II \to III$ is
        only a permitted dynamical track if $\eta_\star \in 0, \eta_{\min,
        III}$.

    \item $III \to I$ --- This track is depicted in \autoref{fig:ad_31}.
        The trajectory encounters the separatrix when $A_I(\eta_\star) +
        A_{II}(\eta_\star) = A_i$, upon which it transitions to a zone I
        trajectory enclosing $A_f = -A_I$.

        This track can only occur if $A_i > A_{II}(\eta_c)$, but is also
        constrained by requiring its $A_i$ is sufficiently small it will
        encounter the separatrix at all (if it is too large, it will never
        encounter the separatrix, which we classify as a $III \to III$
        transition below). This is written $A_i < \max A_I + A_{II} = 4\pi -
        \min A_{III}$. Within these constraints, the zone transition occurs with
        probability
        \begin{equation}
            \Pr_{III \to I} = -\at{\frac{\pd{A_I}{\eta}}{\pd{A_{III}}{\eta}}}
                _{\eta_\star}.
        \end{equation}
        Note that since $\pd{A_I}{\eta} < 0$ always while $\pd{A_{III}}{\eta} >
        0$ for all accessible $\eta_{\star}$, this track is always permitted.

    \item $III \to II \to I$ --- This track is depicted in
        \autoref{fig:ad_321}. The first separatrix encounter occurs at $\eta_1$
        when $A_I(\eta_1) + A_{II}(\eta_1) = A_i$, upon which the trajectory
        moves into zone II enclosing intermediate phase space area $A_m =
        A_{II}(\eta_1)$. Examining \autoref{fig:eq_areas}, there exist values
        $A$ for which $A_{II}(\eta) = A$ has multiple solutions. These are the
        only values for which the $III \to II \to I$ transition is permitted.
        Thus, as $\eta$ continues to decrease, a second $\eta_2$ value exists
        for which $A_m = A_II(\eta_2)$, upon which the trajectory is ejected to
        zone I and $A_f = -A_I(\eta_2)$. The final obliquity again then obeys
        \autoref{eq:qfaf}.

        Similar to the $III \to I$ track, the $III \to II \to I$ track can only
        occur over the same $A_{II}(\eta_c) < A_i < \max A_I + A_{II}$ interval.
        Here, the transition probability for the initial $III \to II$ transition
        is simply
        \begin{equation}
            \Pr_{III \to II} = -\at{\frac{\pd{A_II}{\eta}}{\pd{A_{III}}{\eta}}}
                _{\eta_1}.
        \end{equation}
        It bears noting that this expression is only nonzero for a very small
        fraction of $A_i$ values, since it requires $\pd{A_{II}}{\eta}$ and
        $\pd{A_{III}}{\eta}$ to have different signs; this occurs only if
        $A_{II}\p{\eta_c} < A_i < A_{II, \max}$. Outside of these bounds,
        $\Pr_{III \to II} < 0$ which is interpreted again as a forbidden
        transition.

        Then, once a $III \to II$ transition occurs, the second $II \to I$
        transition occurs for some $\eta_2$ satisfying $A_{II}(\eta_2) =
        A_{II}(\eta_1), \eta_2 < \eta_1$. Graphical inspection shows that
        $\pd{A_{II}}{\eta}$ has the same sign as $\pd{A_{III}}{\eta}$ over all
        possible values, so the second $II \to I$ transition is guaranteed,
        completing the $III \to II \to I$ track.

    \item $III \to III$ --- This track depicts the trivial case where no
        separatrix encounter ever occurs, and $A$ is constant throughout the
        evolution. This constraint equates to $A_i > \max A_I + A_{II}$,
        whereupon no separatrix encounter occurs at all.
\end{itemize}
In summary, the five regimes of track possibilities in increasing order of $A_i$
are:
\begin{itemize}
    \item $II \to III, II \to I$ both possible, $A_i < A_{II}\p{\eta_{\min,
        III}}$.

    \item $II \to I$ only, $A_{II}\p{\eta_{\min, III}} < A_i < A_{II}(\eta_c)$.

    \item $III \to I, III \to II \to I$ both possible, $A_{II}(\eta_c) < A_i <
        A_{II, \max}$.

    \item $III \to I$ only, $A_{II, \max} < A_i < \max A_I + A_{II}$.

    \item $III \to III$ only, $A_i > \max A_I + A_{II}$.
\end{itemize}
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/3testo21.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/3testo21_subplots.png}
    \end{subfigure}
    \caption{Fiducial simulation following the $A_{II} \to A_{I}$ transition.
    Top plot upper panel: plot of $\cos \theta(t)$ (gold) in an example
    simulation. Overlaid are the locations of Cassini State 2 (magenta), and
    upper and lower bounds on the separatrix (black). Note that the trajectory
    successfully tracks CS2 to a misalignment angle $\theta \approx 90^\circ$.
    Vertical green lines denote times portrayed in bottom plot. Top plot lower
    panel: plot of the enclosed separatrix area obtained by integrating the
    simulated trajectory (blue dots) and adiabatic theory (red line). Lower
    plot: snapshots in the $\p{\cos \theta, \phi}$ space of the simulation
    trajectories for the times demarcated in green above, corresponding to the
    start of the simulation, the appearance of the separatrix, two panels
    depicting the separatrix crossing process, and a final snapshot at $\eta =
    10^{-3.5}$. Gold dots denote a full circulation or libration cycle at the
    selected time, while the separatrix at the start/end of the portrayed cycle
    are shown in solid/dashed black lines respectively. Also labeled is CS2
    (magenta). Finally, the enclosed phase space area is shaded in grey ($A >
    0$) and red ($A < 0$).}\label{fig:ad_21}
\end{figure}
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/3testo23.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/3testo23_subplots.png}
    \end{subfigure}
    \caption{Same as \autoref{fig:ad_21} but for the $A_2 \to A_3$ track.
    }\label{fig:ad_23}
\end{figure}
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/3testo31.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/3testo31_subplots.png}
    \end{subfigure}
    \caption{Same as \autoref{fig:ad_21} but for the $A_{III} \to A_I$
    track.}\label{fig:ad_31}
\end{figure}
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/3testo321.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/3testo321_subplots.png}
    \end{subfigure}
    \caption{Same as \autoref{fig:ad_21} but for the $A_{III} \to A_{II} \to A_I$
    track. Two separatrix crossings are shown.}\label{fig:ad_321}
\end{figure}

@@

Next, we will consider the evolution of a system with arbitrary $\theta_{sd, i}$
starting in the $\eta \gg 1$ regime. By evolving such systems until $\eta \ll
1$, we can measure the final $\theta_f$ values and make comparisons with those
obtained via the dynamical tracks above. Observe that when $\eta \gg 1$,
trajectories librate with approximately constant $\theta_{sd}$, meaning they
enclose phase space area
\begin{equation}
    A_i = 2\pi\p{1 - \cos \theta_{sd, i}}.\label{eq:ai_qsd}
\end{equation}
Thus, for a given $\theta_{sd, i}$, we know immediately which tracks are
accessible and with what probability following the results of
\autoref{ss:zone_transitions}.

We run simulations over a ring of initial conditions at each $\theta_{sd, i}$
and measure the final $\theta_{f}$ values. We use \autoref{eq:ai_qsd} to predict
the final obliquities $\theta_f$ according to each of the dynamical tracks. The
agreement of these predictions for $I = 5^\circ, 10^\circ, 20^\circ$ can be seen
in \autoref{fig:ad_ensemble}, \autoref{fig:3_ensemble_10_35}, and
\autoref{fig:3_ensemble_20_35}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_05_35.png}
    \caption{Top: $\theta_{f}\p{\theta_{sd, i}}$, overlaid with semi-analytic
    predictions of the $\theta_{ f}$ for each of the four nontrivial
    dynamical tracks in colored lines. Bottom: Semi-analytic probabilities of
    each of the dynamical tracks for each $\theta_{sd, i}$. The five regimes of
    $\theta_{sd, i}$ values correspond to the five regimes of $A_i$ discussed in
    \autoref{ss:zone_transitions} when using \autoref{eq:ai_qsd} to relate
    $A_i\p{\theta_{sd, i}}$. }\label{fig:ad_ensemble}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_10_35.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $I =
    10^\circ$.}\label{fig:3_ensemble_10_35}
\end{figure}
\begin{figure}
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_20_35.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $I =
    10^\circ$.}\label{fig:3_ensemble_20_35}
\end{figure}

\section{Nonadiabatic Effects}\label{s:nonad}

@@

To illustrate the transition to non-adiabaticity, we present three simulations
with increasingly larger values of $\epsilon$ than \autoref{fig:ad_ensemble}
below in \autoref{fig:3_ensemble_05_25} and \autoref{fig:3_ensemble_05_15}.

@@

Later, these large scatters take on band-like structures that seem to persist
across values of $\theta_{sd, i}$; we attribute these to the separatrix crossing
process becoming sensitive to the \emph{phase} of the final
libration/circulation orbit at separatrix crossing. This is equivalent to
separatrix crossing no longer being significantly slower than orbit timescales
and violating the adiabatic assumption.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_05_25.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $\epsilon = 10^{-2.5}$.
    A larger spread from the dynamical tracks is observed in the
    data.}\label{fig:3_ensemble_05_25}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_05_15.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $\epsilon = 10^{-1.5}$.
    Some small semblance of the evolutionary tracks remains, and the deviations
    appear to have a banded structure. This can be attributed to
    non-adiabaticity ``freezing-in'' the phase of the obliquity variations over
    the final libration/circulation orbit prior to separatrix
    crossing.}\label{fig:3_ensemble_05_15}
\end{figure}

\subsection{Strongly Nonadiabatic Fiducial Trajectory}

@@

The agreement of \autoref{eq:nonad_q_f} at fixed $I$ for varying $\epsilon$ is
shown in \autoref{fig:nonad_3_scan}. Note that $\epsilon \to 0$ recovers the
adiabatic regime. The deviation of $\theta_f$ from the analytical prediction
within the adiabatic regime is indeed due to adiabatic effects becoming
dominant; compare to \autoref{fig:nonad_3_scan_20} where little deviation is
observed until $\theta_f \approx 90^\circ$.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3scan.png}
    \caption{Plot of $\theta_{ f}\p{\theta_{sd, i} = 0}$ as a function of
    $\epsilon$, where $I = 5^\circ$. Overplotted in the red line is
    \autoref{eq:nonad_q_f}, which is in good agreement for $\epsilon \gtrsim
    0.1$ the non-adiabatic regime, while $\theta_{ f} \approx 90^\circ$ in
    the adiabatic regime.}\label{fig:nonad_3_scan}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3scan_20.png}
    \caption{Same as \autoref{fig:nonad_3_scan} but for $I=20^\circ$. Note that
    the analytical prediction is accurate until $\theta_f \approx 90^\circ$,
    which occurs before the transition to the adiabatic
    regime.}\label{fig:nonad_3_scan_20}
\end{figure}

@@

\subsection{Analytic $\theta_f$ for small $\theta_{sd,
i}$}\label{ss:app_transition}

In this section, we will use approximations valid for small $\eta$ to derive
analytic expressions for the final obliquities and associated probabilities for
the $II \to I, II \to III$ tracks that govern behavior at small $\theta_{sd,
i}$.

We first seek a simple parameterization for the separatrix. We first solve for
equilibria of the equation of motion \autoref{eq:dsdt_base} to compute the
coordinates for Cassini State 4:
\begin{equation}
    \cos \theta_4 \approx \frac{\mu \cos I}{1 - \eta \sin I}.
\end{equation}
Note that $\phi_4 = 0$. Then, the separatrix is the level curve of the
Hamiltonian intersecting CS4, so it satisfies $H\p{\theta_{sep}(\phi), \phi} =
H\p{\theta_4, \phi_4}$. This solves to be approximately
\begin{equation}
    \cos \theta_{sep}(\phi) \approx \cos \theta_4 \pm
        \sqrt{2\eta \sin I\p{1 - \cos \phi}}.
\end{equation}
Integration of the phase area enclosed by the two legs of the separatrix then
yields
\begin{equation}
    A_{II}(\eta) \approx 16\sqrt{\eta \sin I}.\label{eq:a_approx}
\end{equation}
This, in conjunction with \autoref{eq:henrard_hop}, is sufficient to compute
$\theta_f\p{\theta_{sd, i}}$ for zone II initial conditions.
